### 2016-07-18 P4-Explore-and-Summarize-Data/submission.Rmd
Update submission.Rmd

Wine Quality by Pratyush Kumar
========================================================
# Description
In this project, we will explore a data set on red wines quality. Our main objective is to explore the chemical properties influences the quality of red wines. This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. The data set is available [here](https://www.google.com/url?q=https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv&sa=D&ust=1468567697749000&usg=AFQjCNE7WYwhUvgZV9QX1BC9Hban9oAmRw) and information about the data set is available [here](https://www.google.com/url?q=https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt&sa=D&ust=1468567697751000&usg=AFQjCNHYOEVrLI2HIxVJ9eXxCflrzbHD2A).

```{r echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r }
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r, Load_the_Data}
rw <- read.csv('wineQualityReds.csv')
```

# Statistical Summary 
### Structure of the dataset
```{r , structure}
str(rw)
```

* There are **1,599** observation with total **13** variables.
* ```X``` is an unique identifier.
* All the variables in the dataset have numeric values except ```X``` and ```quality``` which are integer.

### Statistical summary of the variables in the dataset
```{r, summary}
summary(rw)
```
As we are primarily interested in quality of the red wines, let's see some basic statistics about it
```{r }
summary(rw$quality)
```
* We've found that the quality lies in the range of [3, 8] with mean is equal to 5.636.

# Univariate Plots Section
```{r, Univariate_Plots}
p1 <- qplot(data = rw, fixed.acidity, xlab = 'fixed.acidity (g / dm^3)')
p2 <- qplot(data = rw, volatile.acidity, xlab = 'volatile.acidity (g / dm^3)')
p3 <- qplot(data = rw, citric.acid, xlab = 'citric.acid (g / dm^3)')
p4 <- qplot(data = rw, residual.sugar, xlab = 'residual.sugar (g / dm^3)')
p5 <- qplot(data = rw, chlorides, xlab = 'chlorides (g / dm^3)')
p6 <- qplot(data = rw, free.sulfur.dioxide, xlab = 'free.sulfur.dioxide (mg / dm^3)')
p7 <- qplot(data = rw, total.sulfur.dioxide, xlab = 'total.sulfur.dioxide (mg / dm^3)')
p8 <- qplot(data = rw, density, xlab = 'density (g / dm^3)')
p9 <- qplot(data = rw, pH)
p10 <- qplot(data = rw, sulphates, xlab = 'sulpahtes (g / dm^3)')
p11 <- qplot(data = rw, alcohol, xlab = 'alcohol (% volume)')
p12 <- qplot(data = rw, quality)

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 4)
```

# Univariate Analysis

### Quality of wine
As we can see the entire wine quality are in the range of 3 to 8 with the most common values are 5 and 6 and the least common values are 3, 4, 7, and 8. So, we create another variable **rating** with rate given below.

* 0 - 4 : poor
* 5 - 6 : good
* 7 - 10 : ideal

```{r }
rw$rating <- ifelse(rw$quality < 5, 'poor', ifelse(
  rw$quality < 7, 'good', 'ideal'))

rw$rating <- ordered(rw$rating, levels = c('poor', 'good', 'ideal'))
summary(rw$rating)

ggplot(data = rw, aes(x = rating)) +
  geom_bar()
```

### Calculate total acidity
To calculate sum of all acids in the red wines, we create new variable ```total.acidity```.

```{r }
rw$total.acidity <- rw$fixed.acidity + rw$volatile.acidity + rw$citric.acid

head(rw$total.acidity)
```

### Distribution and Outliers

* ```fixed.acidity```, ```volatile.acidity```, ```sulfur.dioxide```, ```sulphated``` and ```alcohol``` are appeared to be long tailed.
* ```density``` and ```pH``` are normally distributed with few outliers.
* ```residual.sugar``` and ```chlorides``` have extreme outliers.
* ```citric.acid``` contains large number of zero values.

### Plot base 10 logarithmic scale of the long tailed distribution

```{r }
ggplot(data = rw, aes(x = fixed.acidity)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('fixed.acidity (g / dm^3)')

ggplot(data = rw, aes(x = volatile.acidity)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('volatile.acidity (g / dm^3)')

ggplot(data = rw, aes(x = free.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('free.sulfur.dioxide (mg / dm^3)')

ggplot(data = rw, aes(x = total.sulfur.dioxide)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('total.sulfur.dioxide (mg / dm^3)')

ggplot(data = rw, aes(x = sulphates)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('sulphates (g / dm^3)')

ggplot(data = rw, aes(x = alcohol)) +
  geom_histogram() + 
  scale_x_log10() +
  xlab('alcohol (% volume')
```

Taking log_10, we can see that ```fixed.acidity```, ```volatile.acidity```, and ```sulphates``` are normally distributed, with some few outliers.

### Find number of zeroes in ```citric.acid```

```{r }
length(subset(rw, citric.acid == 0)$citric.acid)
```

We found that 132 observations have zero values.

### Find some patterns in ```residual.sugar``` and ```chlorides``` after removing some extreme outliers

```{r }
ggplot(data = rw, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.5, quantile(rw$residual.sugar, 0.95))) +
  xlab('residual.sugar (g / dm^3)')

ggplot(data = rw, aes(x = chlorides)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.04, quantile(rw$chlorides, 0.95))) +
  xlab('chlorides (g / dm^3)')
```

* After removing some extreme outliers, we see ```chlorides``` are now normally distributed
* Since ```residual.sugar``` comes in wide range as, it' is's rare to find wines with less than 1 gm/liter and wines with greater than 45 gm/liter are considered sweet, so the range of 1 - 4 as we found in the plot are ok with some outliers.

# Some Questions
### What is the structure of your dataset?

```{r }
str(rw)
```

### What is/are the main feature(s) of interest in your dataset?
As our main objective is to conclude ```quality```. So, it's the main feature.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
```density``` and ```pH``` are also normally distributed as our new variable ```rating```. So, these two can help support our analysis.

### Did you create any new variables from existing variables in the dataset?

* I created an ordered factor ```rating```, level each variables as 'poor', 'good' and 'ideal'.
* And, ```total.acidity``` to calculate sum of all acids.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

* ```residual.sugar``` and ```chlorides``` contains many outliers but after doing some operations, ```chlorides``` get into normal distribution
* ```citric.acid``` have very large number of zero values but after reading documentation it's fine as it found in small quantities.

# Bivariate Plots Section

### Boxplot of ```quality```

```{r }
quality_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(as.factor(x), y)) +
           geom_boxplot(fill = 'green') +
            xlab('quality') +
            ylab(ylab))
}

grid.arrange(quality_plot(rw$quality, rw$fixed.acidity, 'fixed.acidity (g / dm^3)'),
             quality_plot(rw$quality, rw$volatile.acidity, 
                          'volatile.acidity (g / dm^3)'),
             quality_plot(rw$quality, rw$citric.acid, 'citric.acid (g / dm^3)'),
             quality_plot(rw$quality, rw$residual.sugar, 'residual.sugar (g / dm^3)'),
             quality_plot(rw$quality, rw$chlorides, 'chlorides (g / dm^3)'),
             quality_plot(rw$quality, rw$free.sulfur.dioxide, 
                          'free.sulfur.dioxide (mg / dm^3)'),
             quality_plot(rw$quality, rw$total.sulfur.dioxide, 
                          'total.sulfur.dioxide (mg / dm^3)'),
             quality_plot(rw$quality, rw$density, 
                          'density (g / cm^3)'),
             quality_plot(rw$quality, rw$pH, 'pH'),
             quality_plot(rw$quality, rw$sulphates, 'sulphates (g / dm^3)'),
             quality_plot(rw$quality, rw$alcohol, 'alcohol (% volume)'),
             quality_plot(rw$quality, rw$total.acidity, 'alcohol (% volume)'),
             ncol = 4)
```

### Boxplot of ```rating```

```{r }
rating_plot <- function(x, y, ylab) {
  return (ggplot(data = rw, aes(x, y)) +
           geom_boxplot(fill = 'orange') +
            xlab('rating') +
            ylab(ylab))
}

grid.arrange(rating_plot(rw$rating, rw$fixed.acidity, 'fixed.acidity (g / dm^3)'),
             rating_plot(rw$rating, rw$volatile.acidity, 
                         'volatile.acidity (g / dm^3)'),
             rating_plot(rw$rating, rw$citric.acid, 'citric.acid (g / dm^3)'),
             rating_plot(rw$rating, rw$residual.sugar, 'residual.sugar (g / dm^3)'),
             rating_plot(rw$rating, rw$chlorides, 'chlorides (g / dm^3)'),
             rating_plot(rw$rating, rw$free.sulfur.dioxide, 
                         'free.sulfur.dioxide (mg / dm^3)'),
             rating_plot(rw$rating, rw$total.sulfur.dioxide, 
                         'total.sulfur.dioxide (mg / dm^3)'),
             rating_plot(rw$rating, rw$density, 'density (g / cm^3)'),
             rating_plot(rw$rating, rw$pH, 'pH'),
             rating_plot(rw$rating, rw$sulphates, 'sulphates (g / dm^3)'),
             rating_plot(rw$rating, rw$alcohol, 'alcohol (% volume)'),
             rating_plot(rw$rating, rw$total.acidity, 'alcohol (% volume)'),
             ncol = 4)
```

### Mean of variables for *ideal* rating

```{r}
mean <- c(mean(subset(rw, rating == 'ideal')$fixed.acidity),
          mean(subset(rw, rating == 'ideal')$volatile.acidity),
          mean(subset(rw, rating == 'ideal')$citric.acid),
          mean(subset(rw, rating == 'ideal')$residual.sugar),
          mean(subset(rw, rating == 'ideal')$chlorides),
          mean(subset(rw, rating == 'ideal')$free.sulfur.dioxide),
          mean(subset(rw, rating == 'ideal')$total.sulfur.dioxide),
          mean(subset(rw, rating == 'ideal')$density),
          mean(subset(rw, rating == 'ideal')$pH),
          mean(subset(rw, rating == 'ideal')$sulphates),
          mean(subset(rw, rating == 'ideal')$alcohol),
          mean(subset(rw, rating == 'ideal')$total.acidity))

names(mean) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'residual.sugar', 'chlorides', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'sulphates', 'alcohol', 'total.acidity')

mean
```

### Median of variables for *ideal* rating

```{r}
median <- c(median(subset(rw, rating == 'ideal')$fixed.acidity),
          median(subset(rw, rating == 'ideal')$volatile.acidity),
          median(subset(rw, rating == 'ideal')$citric.acid),
          median(subset(rw, rating == 'ideal')$residual.sugar),
          median(subset(rw, rating == 'ideal')$chlorides),
          median(subset(rw, rating == 'ideal')$free.sulfur.dioxide),
          median(subset(rw, rating == 'ideal')$total.sulfur.dioxide),
          median(subset(rw, rating == 'ideal')$density),
          median(subset(rw, rating == 'ideal')$pH),
          median(subset(rw, rating == 'ideal')$sulphates),
          median(subset(rw, rating == 'ideal')$alcohol),
          median(subset(rw, rating == 'ideal')$total.acidity))

names(median) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'residual.sugar', 'chlorides', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'sulphates', 'alcohol', 'total.acidity')

median
```

# Bivariate Analysis

**poor** rating seems to have following trends:

* lower ```fixed.acidity```, higher ```volatile.acidity``` and lower ```citric.acid```
* lower ```sulfur.dioxide``` and ```sulphates```
* higher ```pH``` and high ```density

**good** rating seems to have following trends:

* low ```fixed.acidity``` and ```volatile.acidity```
* higher ```sulfur.dioxide```
* low ```pH``` and higher ```density```

**ideal** rating seems to have following trends:

* higher ```fixed.acidity``` , lower ```volatile.acidity``` and higher ```citric.acid```
* low ```sulfur.dioxide``` and higher ```sulphates```
* lower ```pH``` and ```density```

### Correlation of variables against quality

```{r }
correlations <- c(cor.test(rw$fixed.acidity, rw$quality)$estimate,
                  cor.test(rw$volatile.acidity, rw$quality)$estimate,
                  cor.test(rw$citric.acid, rw$quality)$estimate,
                  cor.test(rw$residual.sugar, rw$quality)$estimate,
                  cor.test(rw$chlorides, rw$quality)$estimate,
                  cor.test(rw$free.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$total.sulfur.dioxide, rw$quality)$estimate,
                  cor.test(rw$density, rw$quality)$estimate,
                  cor.test(rw$pH, rw$quality)$estimate,
                  cor.test(rw$sulphates, rw$quality)$estimate,
                  cor.test(rw$alcohol, rw$quality)$estimate,
                  cor.test(rw$total.acidity, rw$quality)$estimate)

names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'residual.sugar', 'chlorides', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'sulphates', 'alcohol', 'total.acidity')
correlations
```

**Following variables show strong correlations with ```quality```**

* ```alcohol```
* ```sulphates```
* ```citric.acid```
* ```fixed.acidity```

**Let's see some relationships between some variable and ```total.acidity```**

```{r }
ggplot(data = rw, aes(log10(total.acidity), density)) +
  geom_point(alpha = 1/10) +
  scale_x_continuous() +
  geom_smooth() +
  xlab('log10[total.acidity (g / dm^3)]') +
  ylab('density (g / dm^3)')

ggplot(data = rw, aes(log10(total.acidity), pH)) +
  geom_point(alpha = 1/10) +
  scale_x_continuous() +
  geom_smooth() +
  xlab('log10[total.acidity (g / dm^3)]')

```

We see that there is approx. linear relationship between density and log10(total.acidity), and pH and log10(total.acidity).

# Multivariate Plots Section

```{r , Multivariate_Plots}
ggplot(data = rw, aes(pH, alcohol, color = as.factor(quality))) +
  geom_point() +
  facet_wrap(~rating) +
  ylab('alcohol (% volume)') +
  theme_dark() +
  scale_color_brewer(type = 'seq', palette = 1)

ggplot(data = rw, aes(alcohol, citric.acid, color = as.factor(quality))) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('alcohol (% volume)') +
  ylab('citric.acid (g / dm^3)') +
  scale_color_brewer(type = 'seq', palette = 2) +
  theme_dark()

ggplot(data = rw, aes(alcohol, sulphates, color = as.factor(quality))) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('alcohol (% volume)') +
  ylab('sulphates (g / dm^3)') +
  scale_color_brewer(type = 'seq', palette = 3) +
  theme_dark()

ggplot(data = rw, aes(citric.acid, fixed.acidity, color = as.factor(quality))) +
  geom_point() +
  facet_wrap(~rating) +
  xlab('citric.acid (g / dm^3)') +
  ylab('fixed.acid (g / dm^3)')  +
  scale_color_brewer(type = 'seq', palette = 4) +
  theme_dark()
```

# Multivariate Analysis
I plot graph between four variables ```citric.acid```, ```fixed.acidity```, ```sulphates``` and ```alcohol``` which shown high correlations with ```quality``` and faceted them with ```rating```. I conclude that higher ```citric.acid``` and lower ```fixed.acidity``` yields better wines. Better wines also have higher ```alcohol``` and ```sulphates``` and lower ```pH```.

------

# Final Plots and Summary

### Plot One : Effect of alcohol on wine quality
```{r , Plot_One}
ggplot(data = rw, aes(quality, alcohol, fill = rating)) +
  geom_boxplot() +
  ggtitle('Alcohol Levels on Wine Quality') +
  xlab('Quality') +
  ylab('Alcohol (% volume)')

```

### Description One
As alcohol is highly correlated with the quality, it is better to see its pattern with varying rating. From the above plot, it clearly shows higher % of alcohol yields better wine.

### Plot Two : Effect of acids on wine quality
```{r , Plot_Two}
grid.arrange(ggplot(data = rw, aes(quality, fixed.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, volatile.acidity, fill = rating)) +
               geom_boxplot(),
             ggplot(data = rw, aes(quality, citric.acid, fill = rating)) +
               geom_boxplot(),
             ncol = 2)
```

### Description Two
As more the acidic better is the wine. It would be better to see which acids have more impact on wine quality. Above plot shows, ```fixed.acidity``` and ```citric.acid``` have highly correlated with quality but ```volatile.acidity``` has negative impact on quality.

### Plot Three : Ideal vs poor wine

```{r } 
ggplot(data = subset(rw, rating != 'good'), aes(alcohol, sulphates, color = rating)) +
  geom_point() +
  ggtitle('Alcohol vs Sulphates on Wine Quality') +
  xlab('Sulphates (g / dm^3)') +
  ylab('Alcohol (% volume)')
```

### Description Three
It would be great to see the real pattern between good and bad wines. Above plot differentiate between good and bad wines. It shows higher the % of alcohol and higher the sulphates give better wines.

------

# Reflection
So, there are lots of features on which the wine quality is depend. We do a lot to find the relationship between every variables in the dataset, try to get some linear relation using geom_smooth().  
And, after this EDA, I can conclude that the major factors for better wine quality is alcohol, acidity and sulphates. These features must also be in required content otherwise negative impact will effect the wine quality. Also, we can't be totally sure about quality index also it has been taken by some experts. We've also concluded that there is linear relationship between pH and quality with negative slope.

One thing that is still unclear is the amount of ```residual.sugar```. It contains many outliers, also after doing some operation we get its common range from 1 to 4. But we can't find its amount for ideal wine quality. I think more future research need to be done to find its ideal quantity.